plyranges: a grammar for transforming genomics data

### 2018-03-21 paper.Rmd
rejigged para 1

Genomic data may be naturally represented as sets of pairs of integers 
corresponding to the start and end points of sequences. Further information
such as strandedness and chromosome name may be added to these sets to 
provide biological context. Because of the flexibility of this representation 
supplemental information such as measurements obtained from an experimental 
assay or annotations from a genome database can be joined to their relevant
sequences. In the Bioconductor/R packages `IRanges` and `GenomicRanges` these 
representations have been implemented as a suite of data structures called
_Ranges_ @Lawrence2013-wg; @Huber2015-ei. These data structures cover 
many common data types  encountered in bioinformatics analyses. For
example, a gene can be represented with its 
coordinates, along with its identifier and the identifiers of its exons; or
an RNA-seq experiment may be represented as sets of genes with a matching count 
column.

>>>>

There is a large variety of genomics data available to the modern biologist
from sequencing assays to public collections: annotations of genomic features
such as genes or exons for reference genomes, coverage vectors in the BigWig
format, peaks called from ChIP-Seq experiments and many more.
All of these hetergenous data types can be naturally represented as annotated sets of 
integer pairs, corresponding to the start and end points of sequences, along with 
strandedness and chromosome location to provide biological context. Measurements
or metadata can be included as additional features on their relevant sequences. 
As an example agene can be represented with its coordinates, along with its 
identifier and the identifiers of its exons; or an RNA-seq experiment may be
represented as sets of genes with a matching count column.

In the Bioconductor/R packages `IRanges` and `GenomicRanges` these 
representations have been implemented as a suite of data structures called
_Ranges_ @Lawrence2013-wg; @Huber2015-ei. These data structures cover 
many common data types  encountered in bioinformatics analyses.
